The Companion Mass Distribution to the ELM WD Sample

### 2014-10-02 documents/letter/ms.tex
describing experiments

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. However, in tests below, we generate test data with a variety of mixture component forms. The likelihood with a Gaussian mixture model is
\begin{align}
	h(M_2, m_f, M_1) &= \frac{(M_1+M_2)^{4/3}}{3m_f^{1/3}M_2\sqrt{M_2^2 - (m_f(M_1+M_2)^2)^{2/3}}}\\
	p(m_f \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, m_f, M_1)\\
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD}) \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS})
\end{align}
where $\mathcal{N}$ is the normal distribution with mean $\mu$ and variance $\sigma^2$. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.2~\Msun)^2
\end{align}
given the uncertainties in neutron star mass, birth channels, and mass accretion scenarios \citep{refs} [TODO: Jeff, references]. Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$.

@@

@@

@@

Maximizing the likelihood now requires a numerical technique to probe the parameter space. We have chosen to use a Markov Chain Monte Carlo (MCMC) technique. Because we only have three parameters, we can use a standard Metropolis-Hastings acceptance-rejection technique. Using this technique, we randomly generate new model parameters based on the old model parameters, using a proposal distribution, $K(\mathbf{\Theta '} | \mathbf{\Theta})$. So that the proposal distribution is symmetric ($K(\mathbf{\Theta '} | \mathbf{\Theta})$ = $K(\mathbf{\Theta} | \mathbf{\Theta '})$), we follow the standard prescription and adopt a Gaussian distribution for $K(\mathbf{\Theta '} | \mathbf{\Theta})$, centered on $\Theta$. After a "burn-in" of 500 steps, we use 10 separate chains to determine the posterior probability, $P(\mathbf{\Theta} | \vec{D})$.

To test the performance of our model, we apply it to three separate sets of mock data, each with 100 systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution then applying a randomly drawn inclination angle. We apply our model to the distribution of resulting $M_{2,min}$ and $M_1$ values, attempting to recover the initial model parameters.

>>>>

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. However, in tests below, we generate test data with a variety of mixture component forms. We truncate the distributions using physically motivated bounds so that the white dwarf component is restricted to the range $M_2\in (0.2,1.4)~\Msun$ and the neutron star component is restricted to the range $M_2\in (1,2)~\Msun$. The likelihood with a Gaussian mixture model is
\begin{align}
	h(M_2, m_f, M_1) &= \frac{(M_1+M_2)^{4/3}}{3m_f^{1/3}M_2\sqrt{M_2^2 - (m_f(M_1+M_2)^2)^{2/3}}}\\
	p(m_f \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, m_f, M_1)\\
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD});\,(0.2 < M_2 < 1.4~\Msun) \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS});\,(1.4 < M_2 < 2~\Msun)
\end{align}
where $\mathcal{N}$ is the (truncated, but normalized) normal distribution with mean $\mu$ and variance $\sigma^2$ and the distributions are limited to the ranges specified. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.1~\Msun)^2
\end{align}
given the uncertainties in neutron star mass, birth channels, and mass accretion scenarios \citep{refs} [TODO: Jeff, references]. Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$.

@@

To test the performance of this Gaussian mixture model, we apply it to three separate sets of mock data (described in detail below), each with 100 ``observed'' systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution, then using a randomly drawn inclination angle and primary mass, $M_1$, to compute the mass function, $m_f$. We draw primary masses from a uniform distribution, $M_1 \sim \mathcal{U}(0.2,0.4)~\Msun$. We apply the same Gaussian mixture model to all test datasets to infer the parameters of the white dwarf mixture component and neutron star fraction.

In each experiment below, we use an ensemble Markov Chain Monte Carlo algorithm \citep{goodman10} to draw samples from the posterior distribution, $p(\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS} \given \bs{m}_f, \bs{M}_1)$. The algorithm is implemented in the \texttt{Python} programming language \citep{foremanmackey13} and uses an ensemble of individual ``walkers'' to naturally adapt to the geometry of the parameter-space being explored. We run the walkers for an initial, burn-in period of 1000 steps starting from randomly drawn initial conditions (sampled from the priors in Table~\ref{tbl:parameters}). We then re-initialized the walkers from their positions at the end of this run, throw out the burn-in samples, and run again for 1000 steps. We compute the autocorrelation time for each parameter to confirm they have converged.

@@

As a first test, we generate $M_2$'s by drawing from a single, truncated Gaussian with mean, variance, and bounds
\begin{align}
	\mu_{\rm WD, true} &= 0.7~\Msun\\
	\sigma^2_{\rm WD, true} &= (0.2~\Msun)^2\\
	0.2 < M_2& < 1.4~\Msun.
\end{align}
The distribution of mock observations of the mass function, $m_f$, are shown by the top-left panel of Figure~\ref{}.

@@
